Vigilantism and Institutions:\\ Understanding Attitudes toward Lynching in Brazil

### 2022-12-01 response-reviewers/response-reviewers.tex

\vspace{.5cm}

\noindent Dear Editor and Reviewers,

\vspace{.5cm}

We would like to thank you for the opportunity to revise our manuscript,
``Vigilantism and Institutions: Understanding Attitudes toward Lynching in
Brazil'' (Ms. No. RAP-22-0164). Following the comments and suggestions of the
editors and reviewers, we have thoroughly revised our article and we believe it
has improved significantly as a result.

We were especially committed to discussing the scope conditions of our article
and to clarifying some technical questions about our experiments as highlighted
by Reviewers 1 and 2. We did this by editing the Introduction, adding a new
paragraph to the Conclusion, and including more details about our methodology
in section two and in the Supplementary Materials. We also removed the
significance tests from the manuscript, as per the suggestion of Reviewer 2,
and added a footnote in the Supplementary Materials indicating that we
randomized the order of the experiments to reduce the risk of order effects.
Moreover, we now show how the findings we report in the article compare with
those found in other contexts, as suggested by Reviewer 1. We also discuss the
results of our subgroup analyses in the Supplementary Materials, again
following the recommendation of Reviewer 1. Lastly, we include all references
suggested by Reviewer 1 and 2 in the main body of the manuscript. We are
extremely appreciative of these recommendations and agree that the manuscript
is now stronger and appeals to a broader readership.

@@

\noindent The Authors

\newpage

\section*{Editor Comments and Responses}

\textbf{1)} The editor writes: ``I agree with R2 that the framing of the study
could be improved. You currently cite a number of interesting figures and basic
facts about lynching in Brazil in the start of the conclusion. I think this
would fit better in the introduction. R2 also argues that the reference to the
Blair et al study on greater costs of local violence is imprecise. I am
guessing you mean something along the lines of likely to affect a greater
number directly than terrorism or civil war, which I think is reasonable but
could be said more directly.''

@@

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

We have moved the facts about lynchings in Brazil to the Introduction and we
agree that the section now provides more context for our experiments. The
second paragraph of the Introduction is as follows: ``\textit{One of the most
serious forms of extrajudicial violence is lynching. Lynching can be defined
as ``incidents of physical violence committed by large numbers of private
citizens against one or more individuals accused of having committed a
``criminal'' offense, whether or not this violence resulted in the death of
the victim(s)'' \citep[645]{godoy2004justice}. Although lynchings occur in
more than one hundred countries in all regions of the world
\citep{jung2020lynching,smith2019contradictions}, Latin America has been
particularly affected by a sharp increase in vigilante violence. Lynching
episodes have been reported in Guatemala, Mexico, Peru, Venezuela and other
countries in the region \citep{barbara2015vigilantes, cruz2019determinants,
godoy2004justice}. From 2011 to 2015, Brazil registered about 2,500 lynching
episodes, and 173 people were killed by angry mobs in 2015 alone---nearly one
execution every two days \citep{barbara2015vigilantes, oliveira2016mob}.
According to José de Souza Martins \citeyearpar{martins2015linchamentos}, who
has studied lynchings in Brazil for more than thirty years, these figures are
not only the highest in the country's history, but among the highest in the
world. The people who participate in lynchings are typically young men, but
they also sometimes include teenage women and girls, elderly women, and even
members of the local police \citep{moura2017linchamentos}.}

We have also changed the sentence in which we cite \citet{blair2017predicting}
to make our claims more explicit. The last sentence in the first paragraph now
reads: ``\textit{This omission is significant not only because local violence
is pervasive in weak states, but also because it often comes at a greater
social cost than terrorism or civil wars \citep{blair2017predicting}}.''

\end{quote}

@@

\vspace{.3cm}

\noindent \textbf{2)} The editor writes: ``The manuscript is focused primarily
on the empirical approach, but I agree with R1 that you could say a bit more
about the rationale for the covariates considered, perhaps drawing out why more
``novel'' features such as family or type of crime is likely to be important.''

@@

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

We have followed Reviewer 1's suggestion and have extended section two (``When
Is Lynching Perceived as More Justified?'')  to provide further information
about our choice of covariates. The third paragraph of that section now reads:
``\textit{In addition to our knowledge of the Brazilian case, we also chose our
attributes based partly on existing work on crime, vigilantism, and
extrajudicial violence. From the prison violence literature, we know that the
relative age differences and the genders of perpetrators and victims affect
how it is perceived \citep{fleisher2009myth}. We likewise know that certain
offenses are seen as especially reprehensible \citep{skarbek2014social}, so
we included such offenses, like molestation, alongside less serious offenses.
We include race because, in the United States, violent lynchings have often
been wielded for racist reasons \citep{dray2003hands}. Likewise, we include
residency because extrajudicial violence is often used against people ``who
don't belong'' or are ``outsiders'' in some respect. Finally, in honor
cultures that value retaliatory violence, people believe that victims and the
family of victims have a special right (and often an obligation) to enact
retribution \citep{weiner2013rule}.}

As we show below, we have also included an additional paragraph in the
Conclusion to discuss how our findings relate to the literature on extralegal
violence.

\end{quote}

@@

\vspace{.3cm}

\noindent \textbf{3)} The editor writes: ``On the presentation of the results,
I agree with R2 that the discussion of the results does not need to re-report
all the significance tests, and the full tables could be included in the text
or appendix.''

@@

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}

We have deleted the statistical details from the main text to improve the
readability of the manuscript. All significance tests are now reported only in
the Supplementary Materials.

\end{quote}

@@

\vspace{.3cm}

\noindent \textbf{4)} The editor writes: ``It would be useful to add a bit more
discussion of Brazil in comparative perspective. I wonder if the results for
Brazil are likely to be more general and that the US could be an outlier, for
example.''

\noindent \textbf{Response:} 
\begin{quote}

We have made the changes requested by the editor and Reviewer 1 and our
Conclusion section has been significantly expanded. We are especially grateful
for this suggestion, as it has allowed us to better situate our article in the
broader literate on vigilantism. Paragraph three of the Conclusion, the longest
in the section, is now exclusively dedicated to the scope conditions of our
research. We have linked our findings to important works in the literature and
have also discussed how a culture of honor may explain some of our main
findings.

The paragraph reads: ``\textit{These findings have implications that extend
well beyond Brazil. In particular, our experiments indicate that lynchings
have a crucial gendered aspect, which has also been documented in other
contexts. Using data from 18 Latin American countries,
\citet{nivette2016institutional} also finds that respondents are most likely
to support lynchings when the criminal raped a child, and studies about
lynchings in the American South point out that several episodes resulted from
accusations of sexual assault \citep{jacquet2013giles, smaangs2020race}. We
believe that a culture of honor may explain these results. Individuals in
honor societies view crime as an attack on their personal reputation and, in
turn, are more likely to take revenge to defend their status and that of
people perceived as deserving protection, such as women and children
\citep{nisbett2018culture}. A culture of honor may also explain why
Brazilians see lynchings carried out by the family of the victim as more
justified, as well as refraining from using extralegal violence if it can
trigger vendettas. We expect similar results in societies which share those
cultural norms. Lastly, the fact that race does not appear to be a major
motivation behind lynchings also reflects the experience of places like Haiti
or Southern Africa, where popular violence was mainly driven by other social
factors \citep{berg2011globalizing, jung2020lynching}. Even in the American
South, Whites and Blacks also lynched people of their own race
\citep{beck1997race}. In this respect, our paper also highlights that
lynchings in the Global South may be more strongly connected with the idea of
``popular justice'' than with racial animus \citep{martins2015linchamentos}.}''

\end{quote}

\noindent \textbf{5)} The editor writes: ``Finally, in order to ensure any
delays on a decision on the resubmission I would encourage you to make sure
that the final manuscript follows the journal guidelines,
\url{https://journals.sagepub.com/author-instructions/rap}. For example, the
sections should not be numbered, and the references should have first initial
only rather than full names and books should have location before publisher
(please ensure that all references are complete – Grimm looks incomplete to me,
for example).''

\noindent \textbf{Response:} 
\begin{quote}

We have completed the references according to Research \& Politics' guidelines
and the sections are unnumbered as requested.

\end{quote}

We would like to thank the editor Kristian Skrede Gleditsch for his very
helpful comments and suggestions.

@@

\vspace{.3cm}

We would like to thank the editor Kristian Skrede Gleditsch for his very
helpful comments and suggestions.

\section*{Reviewer 1 Comments and Responses}

\noindent \textbf{1)} Reviewer 1 writes: ``The empirical approach is presented
as offering many benefits---I would also acknowledge the drawbacks. And many of
these results do not seem to square with other parts of the world (which is a
feature and I would ask the authors for a little bit of speculation as to why!)
but I would suggest the authors think about the scope conditions imposed by the
setting and how well they seem to extend (the gendered aspect, yes, the race
aspect, perhaps there are some conditions).''

@@

\vspace{.3cm}

\noindent \textbf{Response:} \begin{quote}

Reviewer 1 makes an excellent point. We agree that we had not properly
explained the scope conditions of our research and, as mentioned above, we have
included a long paragraph in the Conclusion to detail

\end{quote}

@@

\noindent \textbf{2.2)} Reviewer 2 writes: ``Moreover, and more substantially, I am not sure how including studies with several effect sizes in a separate model (p. 7) can solve the violation of non-independent effect sizes, required in meta-analyses. The authors do not report the statistical models they used for solving these violations (I did not find it in the appendix either). There are various ways to address such issues, such as robust variance estimation (see Hedges et al., 2010) or using multi-level meta-analyses (see e.g. application in Matthes et al., 2019).

Furthermore, since there were several studies focusing on the same country (e.g., 14 studies investigating the US), did they use the same samples in their analysis? If so, doesn't it violate the non-independence assumption on its own (as they use the same district ``sample'')? If that is the case, the authors should treat this effect size dependency also for the first model (with the 42 main coefficients). These issues are very important since some of the authors' findings are based only on the ``extended models'' in Table 3 (i.e., those including 142 effect sizes, even from studies reporting several effect sizes related to the same sample). If these findings are based on biased estimates, we should be a lot more cautious interpreting them.''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}
    As Reviewer 2 correctly notes, the coefficient extraction procedure we just presented in the previous response is the main source of effect size dependence in our extended sample analyses. An additional source of dependence is the fact that many studies analyse the same cases, sometimes even using the same data sources. The Reviewer is also correct to point out that separating the effect sizes into a restricted and an extended sample does not properly address the dependence of the coefficients within the extended sample. We followed the Reviewer's suggestion rigorously by employing multilevel random effect models in all of our estimations. A thorough explanation of how we identified, grouped the studies that constitute the new levels, and coded them can be found in section H.1 of the Supplementary Materials. We describe the procedure more succinctly in the Data and Methods section in the main paper (last paragraph on page 7):

@@

\noindent \textbf{4)} Reviewer 2 writes: ``our results suggest that study coefficients are highly sensitive to research design choices'' (p. 13) - this is a rather bold statement (``highly sensitive'') since most differences between study designs are nonsignificant (on Table 3). I would tone down this argument much more and address its limitations in the discussion.''

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}
    We have adapted the language when referring to the effects of estimation methods according to the Reviewer's suggestions throughout the paper. For instance, paragraph 6 of the Introduction (p.3) reads:

    ``\textit{However, when we look only at articles that employ regression discontinuity designs, we see that all four papers included in our sample suggest that a hi}''

@@

\vspace{.3cm}

\noindent \textbf{Response:} 
\begin{quote}
    Following Reviewer 2's recommendations, we have more appropriately framed the paper so that we discuss the institutional moderators and their relationship to the theory across the entire piece. Since we changed our model specifications and added a new moderator, our Results section was also significantly altered. We reduced the Meta-Analysis subsection (3.2) and expanded the Meta-Regressions subsection (3.3) to account for the most relevant and auspicious findings. In this sense, in paragraph 2 of subsection 3.3, we explain to the reader how she should interpret the meta-regression results. We believe this makes our arguments about the meta-regression results more compelling. We also added more substantive insights to accompany the description of meta-regression results in each paragraph of section 3.3.

    We present and partly interpret our results in the Introduction section, reporting back to elements from the literature and to the theory. The last paragraph in the Introduction (p.3) reads:

    ``\textit{The meta-regressions provide additional evidence that our study sample is highly heterogeneous and that effect sizes differ substantially according to study specifications. When using an extended sample of 162 coefficients, we find that unicameralism is associated with higher public spending, as predicted by the ``law of $1/n$''. Since most unicameral legislatures in our sample are local governments (municipalities or districts), this result supports the predictions of Moreover, our meta-regressions indicate that larger upper chambers spend more in terms of per capita expenditure than lower chambers, a result that also appears in the binomial tests. Overall, non-majoritarian voting systems seem to decrease government spending, following the idea that the $1/n$ effect grows weaker as the empirical cases distance from the original definition of the law. Finally, meta-regression results confirm that regression discontinuity designs reduce public spending estimates.}''

### 2022-12-01 article/article.tex
add response, edit manuscript

These findings have implications that extend well beyond Brazil. In particular,
our experiments indicate that lynchings have a crucial gendered aspect, which
has also been documented in other contexts. Using data from 18 Latin American
countries, \citet{nivette2016institutional} also finds that respondents are
most likely to support lynchings when the criminal raped a child, and studies
about lynchings in the American South point out that several episodes resulted
from accusations of sexual assault \citep{jacquet2013giles, smaangs2020race}.
We believe that a culture of honor may explain these results. Individuals in
honor societies view crime as an attack on their personal reputation and, in
turn, are more likely to take revenge to defend their status and that of people
perceived as deserving protection, such as women and children
\citep{nisbett2018culture}. A culture of honor may also explain why Brazilians
see lynchings carried out by the family of the victim as more justified, as
well as refraining from using extralegal violence if it can trigger vendettas.
We expect similar results in societies which share those cultural norms.
Lastly, the fact that race does not appear to be a major motivation behind
lynchings also reflects the experience of places like Haiti or Southern Africa,
where popular violence was mainly driven by other social factors
\citep{berg2011globalizing, jung2020lynching}. Even in the American South, it
Whites and Blacks also lynched people of their own race \citep{beck1997race}.
In this respect, our paper also highlights that lynchings in the Global South
may be more strongly connected with ideas of ``popular justice'' than racial
animus \citep{martins2015linchamentos}.

>>>>

These findings have implications that extend well beyond Brazil. In particular,
our experiments indicate that lynchings have a crucial gendered aspect, which
has also been documented in other contexts. Using data from 18 Latin American
countries, \citet{nivette2016institutional} also finds that respondents are
most likely to support lynchings when the criminal raped a child, and studies
about lynchings in the American South point out that several episodes resulted
from accusations of sexual assault \citep{jacquet2013giles, smaangs2020race}.
We believe that a culture of honor may explain these results. Individuals in
honor societies view crime as an attack on their personal reputation and, in
turn, are more likely to take revenge to defend their status and that of people
perceived as deserving protection, such as women and children
\citep{nisbett2018culture}. A culture of honor may also explain why Brazilians
see lynchings carried out by the family of the victim as more justified, as
well as refraining from using extralegal violence if it can trigger vendettas.
We expect similar results in societies which share those cultural norms.
Lastly, the fact that race does not appear to be a major motivation behind
lynchings also reflects the experience of places like Haiti or Southern Africa,
where popular violence was mainly driven by other social factors
\citep{berg2011globalizing, jung2020lynching}. Even in the American South,
Whites and Blacks also lynched people of their own race \citep{beck1997race}.
In this respect, our paper also highlights that lynchings in the Global South
may be more strongly connected with the idea of ``popular justice'' than with
racial animus \citep{martins2015linchamentos}.
